<br>

### Taxonomic Phage classification {.tabset}

<br>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
suppressMessages(library(DT))
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(ggplot2))

# Load data paths
tax_table_sourmash <- "#SOURMASHENV#"
tax_table_genomad <- "#GENOMADENV#"
sample_name <- "#NAMEENV#"

# Helper to load and clean
load_data <- function(path) {
  if (file.exists(path) && file.size(path) > 10) { # check if it exists and has content
    df <- tryCatch({
        read.table(file = path, sep = '\t', header = TRUE, stringsAsFactors = FALSE, check.names = FALSE)
    }, error = function(e) NULL)
    return(df)
  }
  return(NULL)
}

sm_data <- load_data(tax_table_sourmash)
gn_data <- load_data(tax_table_genomad)

# Standardize column names
standardize_cols <- function(df, tool) {
    if (is.null(df)) return(NULL)
    
    # Tool-specific renames
    if (tool == "GeNomad") {
        # GeNomad often uses seq_name
        if ("seq_name" %in% colnames(df)) df <- df %>% rename(Contig = seq_name)
        # Map lineage to taxonomy
        if ("lineage" %in% colnames(df)) df <- df %>% rename(taxonomy = lineage)
        
        # Ensure taxonomy and agreement exist for selection
        if (!"taxonomy" %in% colnames(df)) df$taxonomy <- NA
        if (!"agreement" %in% colnames(df)) df$agreement <- NA
    }
    
    if (tool == "Sourmash") {
        # Sourmash headers from our process are Contig, Similarity, Taxonomy...
        if (!"Taxonomy" %in% colnames(df)) df$Taxonomy <- NA
        if (!"Similarity" %in% colnames(df)) df$Similarity <- NA
    }
    
    # Ensure Contig column exists
    if (!"Contig" %in% colnames(df)) return(NULL)
    
    return(df)
}

sm_data <- standardize_cols(sm_data, "Sourmash")
gn_data <- standardize_cols(gn_data, "GeNomad")

# Merge for comparison
comparison_ready <- FALSE
if (!is.null(sm_data) && !is.null(gn_data)) {
    
    sm_sub <- sm_data %>% 
        select(any_of(c("Contig", "Taxonomy", "Similarity"))) %>%
        rename(Taxonomy_Sourmash = Taxonomy, Similarity_Sourmash = Similarity)
        
    gn_sub <- gn_data %>% 
        select(any_of(c("Contig", "taxonomy", "agreement"))) %>%
        rename(Taxonomy_GeNomad = taxonomy, Agreement_GeNomad = agreement)

    comparison_data <- full_join(sm_sub, gn_sub, by = "Contig") %>%
        mutate(
            Match_Status = case_when(
                is.na(Taxonomy_Sourmash) & !is.na(Taxonomy_GeNomad) ~ "GeNomad Only",
                !is.na(Taxonomy_Sourmash) & is.na(Taxonomy_GeNomad) ~ "Sourmash Only",
                (Taxonomy_Sourmash == Taxonomy_GeNomad) ~ "Perfect Match",
                TRUE ~ "Both (Check Consensus)"
            )
        )
    comparison_ready <- (nrow(comparison_data) > 0)
}
```

#### Comparison Overview

<div class = "blue" style="background-color:#f8f9fa; border-left: 6px solid #007bff; padding: 15px; margin-bottom: 20px;">
**How to interpret this comparison:**  
*   **Sourmash** uses k-mer based matching against a reference database. It is highly accurate for known phages but may miss novel ones.  
*   **GeNomad** uses marker genes and neural networks. It is better at classifying novel phages that aren't exact database matches.  
*   **Consensus**: When both tools agree, you can have high confidence in the taxonomic assignment.
</div>

```{r, echo=FALSE, fig.width=10, fig.height=4}
if (comparison_ready) {
    status_summary <- comparison_data %>% 
        group_by(Match_Status) %>% 
        summarise(Count = n())
    
    ggplot(status_summary, aes(x = Match_Status, y = Count, fill = Match_Status)) +
        geom_bar(stat = "identity") +
        theme_minimal() +
        labs(title = paste("Classification Agreement for", sample_name),
             x = "Agreement Level", y = "Number of Contigs") +
        scale_fill_manual(values = c("Perfect Match" = "#28a745", "Both (Check Consensus)" = "#ffc107", 
                                    "Sourmash Only" = "#17a2b8", "GeNomad Only" = "#6610f2")) +
        theme(legend.position = "none")
} else {
    print("Insufficient data for side-by-side comparison.")
}
```

<br>

**Table 1: Combined Taxonomic Predictions**

```{r, echo=FALSE}
if (comparison_ready) {
    # Identify numeric columns for rounding
    num_cols <- names(comparison_data)[sapply(comparison_data, is.numeric)]
    
    datatable(comparison_data, rownames = FALSE, class = 'hover stripe cell-border', extensions = 'Buttons',
              options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'colvis'), pageLength = 10)) %>%
    formatRound(columns = num_cols, digits = 2) %>%
    formatStyle('Match_Status',
                backgroundColor = styleEqual(c('Perfect Match', 'Both (Check Consensus)', 'GeNomad Only', 'Sourmash Only'),
                                            c('#d4edda', '#fff3cd', '#e2d9f3', '#d1ecf1')))
}
```

#### Sourmash (Raw)

**Tab. 2:**  Detailed taxonomic classification based on Sourmash exact k-mer matching.

```{r, echo=FALSE}
if (!is.null(sm_data)) {
    num_cols_sm <- names(sm_data)[sapply(sm_data, is.numeric)]
    datatable(sm_data, rownames = FALSE, class = 'hover', extensions  = 'Buttons', 
              options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'))) %>%
    formatRound(columns = num_cols_sm, digits = 2)
} else {
    print("No Sourmash taxonomy results available.")
}
```

#### GeNomad (Raw)

**Tab. 3:** Detailed taxonomic classification based on GeNomad's neural network and marker gene approach.

```{r, echo=FALSE}
if (!is.null(gn_data)) {
    num_cols_gn <- names(gn_data)[sapply(gn_data, is.numeric)]
    datatable(gn_data, rownames = FALSE, class = 'hover', extensions  = 'Buttons', 
              options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'))) %>%
    formatRound(columns = num_cols_gn, digits = 2)
} else {
    print("No GeNomad taxonomy results available.")
}
```

<a href="#top">Back to top</a>